---
title: "Project on Chicago"
output: word_document
---

## Prince Agyabeng


Loading the Libraries
```{r}
library(tidyverse)
library(GGally) 
library(ggcorrplot)
library(mice) 
library(VIM) 
library(lubridate)
library(ggcorrplot)
library(MASS) 
library(leaps)
library(caret) 
library(rpart) 
library(RColorBrewer) 
library(rattle) 
library(ranger)
library(nnet)
```


Reading the data
```{r}
chicago <- read_csv("chicago.csv")
```


Examining the struture and summary of the data
```{r}
str(chicago)
summary(chicago)
```

Data cleaning and Preparations
Delete the columns below
```{r}
chicago = chicago %>% dplyr::select(-ID) 
chicago = chicago %>% dplyr::select(-`Case Number`)
chicago = chicago %>% dplyr::select(-`Updated On`)
chicago = chicago %>% dplyr::select(-`X Coordinate`)
chicago = chicago %>% dplyr::select(-`Y Coordinate`)
chicago = chicago %>% dplyr::select(-`Location`)
str(chicago)

```


Converting the date
```{r}
chicago = chicago %>% mutate(Date = mdy_hms(Date))
str(chicago)
```

## Preparing the Data

Converting the Month
```{r}
chicago = chicago %>% mutate(Month = month(Date))
str(chicago)
```



Converting the hour
```{r}
chicago = chicago %>% mutate(Hour = hour(Date))
str(chicago)
```



Factor conversion and recoding  
```{r}
chicago = chicago %>% mutate(Year = as_factor(as.character(Year))) %>%
mutate(Year = fct_recode(Year))

chicago = chicago %>% mutate(Month = as_factor(as.character(Month))) %>%
mutate(Month = fct_recode(Month,
 "Jan" = "1",
"Feb" = "2",
"March" = "3",
"April" = "4",
"May" = "5",
"June" = "6",
"July" = "7",
"Aug" = "8",
"Sept" = "9",
"Oct" = "10",
"Nov" = "11",
"Dec" = "12"))

chicago = chicago %>% mutate(Hour = as_factor(as.character(Hour))) %>%
mutate(Hour = fct_recode(Hour))

chicago = chicago %>% mutate(Arrest = as_factor(as.character(Arrest))) %>%
mutate(Arrest = fct_recode(Arrest,
"Yes" = "TRUE",
"No" = "FALSE"))

chicago = chicago %>% mutate(Domestic = as_factor(as.character(Domestic))) %>%
mutate(Domestic = fct_recode(Domestic,
"Domestic_violence" = "TRUE",
"Non_Domestic_violence" = "FALSE"))

chicago = chicago %>% mutate(chicago = as_factor(as.character(`Primary Type`))) %>%
mutate(`Primary Type` = fct_recode(`Primary Type`))

chicago = chicago %>% mutate(Description = as_factor(as.character(Description))) %>%
mutate(Description = fct_recode(Description))

chicago = chicago %>% mutate(`Location Description` = as_factor(as.character(`Location Description`))) %>%
mutate(`Location Description` = fct_recode(`Location Description`))

chicago = chicago %>% mutate(Domestic = as_factor(as.character(Domestic))) %>%
mutate(Domestic = fct_recode(Domestic))

chicago = chicago %>% mutate(District = as_factor(as.character(District))) %>%
mutate(District = fct_recode(District))

chicago = chicago %>% mutate(Ward = as_factor(as.character(Ward))) %>%
mutate(Ward = fct_recode(Ward))

chicago = chicago %>% mutate(`Community Area` = as_factor(as.character(`Community Area`))) %>%
mutate(`Community Area` = fct_recode(`Community Area`))
str(chicago)

summary(chicago)

```

Viewing missing valuse
```{r}
vim_plot = aggr(chicago, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```
Looking at the graph, we need to remov the latitude and longtitude variables as we missing many vaules.


```{r}
chicago = chicago %>% dplyr::select(-Latitude)
chicago = chicago %>% dplyr::select(-Longitude)
str(chicago)
```


Viewing additional values
```{r}
vim_plot = aggr(chicago, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```


Deleting rows of missingnes
```{r}
chicago = chicago %>% drop_na()
summary(chicago)
```


viewing the data
```{r}
vim_plot = aggr(chicago, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```



### Data Exploration
Begin exploring the data by looking at the count of our response variable only
```{r}
ggplot(chicago, aes(x=Arrest, fill = Arrest)) + geom_bar() + theme_bw()
```


`Data visualization
```{r}
ggplot(chicago, aes(x=`Primary Type`, fill = Arrest)) + geom_bar() + theme_bw()
```

There's seems to be relationship between Primary Type  and Arrest. 


```{r}
ggplot(chicago, aes(x=Domestic, fill = Arrest)) + geom_bar() + theme_bw()
```
There's no relationship between Domestic and Arrest as issues without domestic violence leads to more arrest


```{r}
ggplot(chicago, aes(x=Month, fill = Arrest)) + geom_bar() + theme_bw()
```
From the graph there's a strong relationship between Month and Arrest. In the summer there's seem to be more arrest than the other months.


```{r}
ggplot(chicago, aes(x=District, fill = Arrest)) + geom_bar() + theme_bw()
```



## Building Predictive model

Selecting important variables
```{r}
chicago2 = chicago %>% dplyr::select("Arrest", "Description", "Domestic", "Ward", "Month", "Year", "Location Description", "District")
```



Splitting the data
```{r}
set.seed(123)
train.rows = createDataPartition(y = chicago$Arrest, p=0.7, list = FALSE)
train = chicago2[train.rows,] 
test = chicago2[-train.rows,]
```


Visualize using the training set (looking at relationship between SeriousDlqin2yrs and the other variables).  
```{r}
ggplot(train,aes(x=Month, fill = Arrest)) + geom_bar()

```


Finding more relationhips
```{r}
ggplot(train, aes(Month,Arrest))+
  geom_boxplot(aes(color = Month, fill=Arrest))

```


Let's build a model with Month
```{r}
mod1 = glm(Arrest ~ Month , train, family = "binomial")
summary(mod1)
```
Note the AIC of this model (a measure of model quality) is 186,749. We can use this value to compare this model to others. Smaller AIC is better.  

##Building Classification
For Further Analysis, let's build Classification tree
```{r}
tree1 = rpart(Arrest ~., train, method="class")
fancyRpartPlot(tree1)

```


```{r}
printcp(tree1)
plotcp(tree1)
```



Prediction on the training set
```{r}
treepred = predict(tree1, train, type = "class")
head(treepred)
```


A look at the Accuracy on the training set
```{r}
confusionMatrix(treepred,train$Arrest,positive="Yes")
```


Prediction on the test set
```{r}
treepred_test = predict(tree1, test, type = "class")
head(treepred_test)
```


Confusion Matrix on the test set
```{r}
confusionMatrix(treepred_test,test$Arrest,positive="Yes")
```


###Building Random Forest Model with Caret
Building a Random Forest
```{r}
fit_control = trainControl(method = "cv",  
                           number = 5) 


set.seed(123)  
rf_fit = train(Arrest~Month,train,
                 method = "ranger", 
                 importance = "permutation",
                 trControl = fit_control,
               num.trees = 10)
```


Checking the Random Forest
```{r}
varImp(rf_fit)
rf_fit
```


Prediction of Random Forest
```{r}
predRF = predict.train(rf_fit, train)
head(predRF)
```


Accuracy for Random Forest
```{r warning=TRUE}
confusionMatrix(predRF, train$Arrest, positive = "Yes")
```



### Building Neural Network
```{r}
start_time = Sys.time() 
fitControl = trainControl(method = "cv", 
                           number = 5)

nnetGrid <-  expand.grid(size = 6, decay = 0.1)

set.seed(1234)
nnetBasic = train(Arrest~Month,train,
                 method = "nnet",
                 tuneGrid = nnetGrid,
                 trControl = fitControl,
                 verbose = FALSE,
                 trace = FALSE)

end_time = Sys.time()
end_time-start_time
```


```{r}
nnetBasic
```


Prediction and Confustion matrix Neural Network
```{r}
predNetBasic = predict(nnetBasic, train)

confusionMatrix(predNetBasic, train$Arrest, positive = "Yes")
```


Making Prediction on the test data
```{r}
predNetBasic = predict(nnetBasic, test)

confusionMatrix(predNetBasic, test$Arrest, positive = "Yes")
```



