---
title: "Correlation and Simple Linear Regression Assignment"
output: word_document
---

## Prince Agyabeng


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries needed
```{r}
library(tidyverse)
library(GGally)
library(car)
library(lmtest)
```

Task 1
Describing the Dataset
There are 153 observations made up of 6 variables.Also the data is missing vaues for certain columns. I think Ozone will be the response variable
```{r}
air = airquality
str(airquality)
summary(airquality)

```

Task 2
Removing of missing data rows  
There are now 111 observation and 6 varibles of the new data air2
```{r}
air2 = airquality %>% drop_na()
str(air2)

```
Task 3
A look at ggpairs plot for visualization and correlation.
The best variable to predict Ozone appears to be Temp (correlation = 0.699

```{r}
ggpairs(air2)
ggcorr(air2, label = TRUE)
```
Which variable is most strongly correlated with the “Ozone” variable? Temp
Which variable is least strongly correlated with the “Ozone” variable? Month

Task 4
Finding the relationship between Tempe and Ozone
From the scatterplot, we can validate that there's a strong positive relationship between Temp and Ozone.
```{r}
ggplot(air2, aes(x=Temp, y = Ozone)) + geom_point()
```

Task 5
Build a regression model1 using Temp to predict Ozone
```{r}
model1 = lm(Ozone ~ Temp, data = air2)
summary(model1)

confint(model1)
```
The above model1 is significant as a unit change in Temp -147.64 change in the ozone layer. R2 is 0.488 which is good as as nearly half of the model can explain the data. The Temp variable is significant (p-value < 0.05) and it has explanatory power. The range of the slope falls between -184.818 and -110.473

Task 6
Plot the model 
It seems the model is good as the graph indicate a strong relationship between Temp an Ozone
```{r}
ggplot(air2, aes(x=Temp, y = Ozone)) + geom_point()+
  geom_smooth(method= "lm", se = FALSE)
```

Task 7
Using the predict function
```{r}
testdata = data.frame(Temp = c(80))
predict(model1, newdata = testdata, interval = "predict")
```

Task 8
**Assumption 1** The predictor and response variable mmust have a linear relationship. We can see the graph that there's a linear relationship between Tem and Zone
```{r}
ggplot(air2, aes(x=Temp,y=Ozone)) + geom_point() + geom_smooth(method="lm",se=FALSE, color="red") + theme_bw()

```

**Assumption 2** Model errors (residuals) are independent  
```{r}
dwtest(model1)
```
We fail to reject the null hypothesis with a p-value greater than 0.05. This suggests that the residuals are likely independent.

**Assumption 3** Model residuals exhibit constant variance
```{r}
air2 = air2 %>% mutate(resid1 = model1$residuals)

ggplot(air2,aes(x=Temp,y=resid1)) + geom_point() + theme_bw()

```
From the graph the model residual does not exhibit constant variance thus violating this assumption.

**Assumption 4** Model residuals are Normally-distributed
```{r}
ggplot(air2,aes(x=resid1)) + geom_histogram() + theme_bw()

##For a second check
qqPlot(model1)

```
From the histogram above, The residuals histogram is not normally distributed, thus violating the normal distribution assumption.

Task 9
The model in Task 5 can be used to predict the effects of Temp on the Ozone layer. Any unit of increase of Temp has a negative influence on the Ozone. when using this model, one must understand that the modle violates the model residual assumption and the normal distribution assumption.
